---
title: "P8105_HW2"
author: Yijia Chen
date: 2021-10-05
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1
## Problem 1 (a)

Read and clean the Mr. Trash Wheel sheet:

1. specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
2. use reasonable variable names
3. omit rows that do not include dumpster-specific data
4. round the number of sports balls to the nearest integer

```{r message = FALSE}
trashw_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N534") %>%
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls))
```

## Problem 1 (b)

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable.

```{r message = FALSE}

pdata_2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>%   janitor::clean_names() %>% 
  mutate(year = 2019)

pdata_2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%   janitor::clean_names() %>% 
  mutate(year = 2018)

pdata_df = 
  bind_rows(pdata_2018_df, pdata_2019_df) %>% 
  mutate(month = month.abb[month]) %>% 
  rename(total_precipitation = total) %>% 
  relocate(year)

```

For trashw_df, the number of observation is `r nrow(trashw_df)`. All variables in this dataset are `r names(trashw_df)`. Some key variable examples are "weight_tons" with a mean of `r mean(pull(trashw_df, weight_tons))` and variance as `r var(pull(trashw_df, weight_tons))`, "volume_cubic_yards" with a mean of `r mean(pull(trashw_df, volume_cubic_yards))` and variance as `r var(pull(trashw_df, volume_cubic_yards))`, and "plastic_bottles" with a mean of `r mean(pull(trashw_df, plastic_bottles))` and variance as `r var(pull(trashw_df, plastic_bottles))`. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trashw_df, year == 2019), sports_balls))`.

For pdata_df, the number of observation is `r nrow(pdata_df)`. All variables in this dataset are `r names(pdata_df)`. "Total_precipitation" is the key variable here with a mean of `r mean(pull(pdata_df, total_precipitation))` and variance of `r var(pull(pdata_df, total_precipitation))`. The total precipitation in 2018 is `r sum(pull(pdata_2018_df, total))`.


# Problem 2

First, clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day; replace month number with month name; create a president variable taking values gop and dem, and remove prez_dem and prez_gop; and remove the day variable.

```{r message = FALSE}
pols_df = 
  read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(month = month.abb[month]) %>% 
  pivot_longer(
    c(prez_gop,prez_dem),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "presidency"
  ) %>% 
  filter(presidency != 0) %>% 
  select(-c(day, presidency)) %>% 
  relocate(year, month)
```

Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r message = FALSE}
snp_df = 
  read.csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(
    year = as.numeric(year),
    year = ifelse(year > 2021, year - 100, year),
    month = month.abb[month]
    ) %>% 
  rename(sp_index = close) %>% 
  select(-day)
```

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

```{r message = FALSE}
unemploy_df = 
  read.csv("data/unemployment.csv") %>% 
  pivot_longer(
    c(Jan:Dec),
    names_to = "month",
    values_to = "percent_unemploy"
  ) %>% 
  janitor::clean_names()
```

Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r message = FALSE}

union2_df = 
  full_join(pols_df, snp_df, by = c("year", "month"))

union3_df = 
  full_join(union2_df, unemploy_df, by = c("year", "month"))

```

For "pols_df", it contains `r names(pols_df)`, which represents the number of governors, senators, representatives for two parties (gop = republican, dem = democratic) on the associated date. It also indicate which party the precident belong to. 

For "snp_df", it contains `r names(snp_df)`, which means the closing values of the S&P stock index for the given time. 

For "unemploy_df", it contains `r names(unemploy_df)`, which represents the percentage of unemployment for the given time.

Resulting dataset combined three above datasets together. The dimension of "union3_df" is `r dim(union3_df)` indicating rows and columns respectively. The range of years is [`r range(pull(union3_df, year))`]. Key variables are `r names(union3_df)`. This is a comprehensive version of the three datasets above.


# Problem 3
## Problem 3 (a)

Load and tidy the data. Note that, although these data may seem fairly well formatted initially, the names of a categorical predictor and the case structure of string variables changed over time; you’ll need to address this in your data cleaning. Also, some rows seem duplicated, and these will need to be removed.

```{r message = FALSE}
pbname_df = 
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name),
         ethnicity = replace(ethnicity, ethnicity == "ASIAN AND PACI", "ASIAN AND PACIFIC ISLANDER"),
         ethnicity = replace(ethnicity, ethnicity == "BLACK NON HISP", "BLACK NON HISPANIC"),
         ethnicity = replace(ethnicity, ethnicity == "WHITE NON HISP", "WHITE NON HISPANIC")) %>% 
  distinct()
```

## Problem 3 (b)

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time; this should have rows for ethnicities and columns for year.

```{r message = FALSE}
olivia_df = 
  filter(
    pbname_df,
    gender == "FEMALE",
    childs_first_name == "olivia"
  ) %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth", 
    values_from = "rank"
  ) %>% 
  knitr::kable(caption = "childs_name_olivia_rank_overtime_female")

olivia_df
```

## Problem 3 (c)

Produce a similar table showing the most popular name among male children over time.

```{r message = FALSE}
popular_df = 
  select(pbname_df, -count) %>% 
  filter(
    gender == "MALE",
    rank == "1"
  ) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  ) %>% 
  select(-c(rank, gender)) %>% 
  knitr::kable(caption = "most_popular_childs_name_overtime_male")

popular_df
```

## Problem 3 (d)

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r message = FALSE}
whitem_df = 
  filter(
    pbname_df,
    gender == "MALE",
    ethnicity == "WHITE NON HISPANIC",
    year_of_birth == 2016
  ) %>% 
  ggplot(aes(x = rank, y = count)) + 
  geom_point() + 
  geom_text(aes(label = childs_first_name), hjust = 0, vjust = 0) + 
  ggtitle("white_male_childs_name_rank_2016")

whitem_df
```

