---
title: "P8105_HW2"
author: Yijia Chen
date: 2021-10-05
output: github_document
---

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1
## Problem 1 (1)

Read and clean the Mr. Trash Wheel sheet:

1. specify the sheet in the Excel file and to omit non-data entries (rows with notes / figures; columns containing notes) using arguments in read_excel
2. use reasonable variable names
3. omit rows that do not include dumpster-specific data
4. round the number of sports balls to the nearest integer

```{r}

trashw_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = round(sports_balls))
  
```

## Problem 1 (2)

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable.

```{r}

pdata_2019_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>%   janitor::clean_names() %>% 
  mutate(year = 2019)

pdata_2018_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%   janitor::clean_names() %>% 
  mutate(year = 2018)

pdata_df = 
  bind_rows(pdata_2018_df, pdata_2019_df) %>% 
  mutate(month = month.abb[month]) %>% 
  rename(total_precipitation = total) %>% 
  relocate(year)
  
```

